%% Boilerplate FSF Copyright assignment form.
%% $Id$

\documentclass[a4paper,12pt]{article}

\title{\large\sffamily\bfseries
ASSIGNMENT OF COPYRIGHT IN REVISED FILES OF FSF SOFTWARE PROGRAMS}

\addtolength{\parskip}{1em}

\begin{document} 
 
\maketitle

 For \$1 and other good and valuable consideration received from 
 Free Software Foundation, a not-for-profit corporation of the 
 state of Massachusetts, hereinafter ``FSF'', the undersigned 
 ``Developer'' does hereby agree as follows:
 
 1.(a) Developer hereby agrees to assign, and does hereby assign, to FSF
 Developer's copyright in changes and/or enhancements to programs owned
 by FSF (through authorship by or for FSF or via donations or other
 grants to FSF).  These changes and/or enhancements are herein called
 the Works.
 
 (b) The assignment of par. 1(a) above applies to past and future works
 of Developer within the defined `works' category and also includes
 works based on the works.  FSF has all the rights of a copyright owner
 in the assigned copyrights, subject to the license grantback to
 developer stated below, including the right to enforce the copyright
 in aid of the free software purposes of this agreement, and the right
 to use, license and distribute the works, or works based on the works,
 with the program enhanced thereby or as stand-alone modules, all as
 made or acquired by developer or in modified form.
 
 (c) For the purposes of this contract, a work ``based on the Works''
 means any work that in whole or in part incorporates or is derived
 from all or part of the Works.
 
 (d) FSF agrees to grant back to Developer, and does hereby grant,
 non-exclusive, royalty-free and non-cancellable rights to use the
 Works (i.e., Developer's changes and/or enhancements, not the programs
 that they enhance), as Developer sees fit; this grant back does not
 limit FSF's rights and public rights acquired through this agreement.
 
 2.(a) Developer further agrees to use its best efforts to cause any
 persons developing on behalf of Developer any changes and/or
 enhancements to programs owned by FSF to enter into a
 work-made-for-hire relation with Developer, so that copyright on the
 modifications or new files shall consequently be assigned from
 Developer to FSF following par. 1 above, to the end that FSF may
 encourage widespread free use of its programs without encumbrance by
 proprietary claims or a possibility of such claims.
 
 (b) Where work-made-for-hire status is unobtainable, Developer shall
 obtain assignment of copyright by such person(s) running to FSF or
 Developer for the same purpose.
 
 (c) Developer will report occasionally, on its initiative and whenever
 requested by FSF, the changes and/or enhancements it makes, along with
 Developer's origin of copyright--by work-made-for-hire or assignment
 or both--and to the extent known to Developer any outstanding rights
 or claims of rights, by any person, that might be adverse to the
 rights of Developer or FSF or to such purpose.
 
 3.(a) The parties shall execute such documents and undertake such acts
 as may be requested by the other to implement the letter and spirit of
 this agreement and any conveyance stated herein, subject to a
 limitation of reasonableness of such request with attention to cost
 and time burdens imposed thereby.
 
 (b) The undertakings in par. 3(a), above, are without prejudice to the
 conveyances made through this assignment document.  The intention of
 FSF and Developer is that this assignment document will be
 supplemented rarely, if at all, by other documents.
 
 4. FSF agrees that all distribution of the Works, or of any work
 ``based on the Works'', that takes place under the control of FSF or its
 agents or assignees, shall be on terms that explicitly and perpetually
 permit anyone possessing a copy of the work to which the terms apply,
 and possessing accurate notice of these terms, to redistribute copies
 of the work to anyone on the same terms.  These terms shall not
 restrict which members of the public copies may be distributed to.
 These terms shall not require a member of the public to pay any
 royalty to FSF or to anyone else for any permitted use of the work
 they apply to, or to communicate with FSF or its agents or assignees
 in any way either when redistribution is performed or on any other
 occasion.
 
 5. FSF agrees that any program ``based on the Works'' offered to the
 public by FSF or its agents or assignees shall be offered in the form
 of machine-readable source code, in addition to any other forms of
 FSF's choosing.  However, FSF is free to choose at its convenience the
 media of distribution for machine-readable source code, and may charge
 a fee for providing these works to the public.
 
 6. Developer hereby agrees that if it has or acquires hereafter any
 patent or interface copyright or other intellectual property interest
 dominating the Works, or the program enhanced by the Works, or use of
 that program or of the Works, such dominating interest will not be
 used to undermine the effect of this assignment, i.e. the Foundation
 and the general public will be licensed to use, in that program or
 programs and their derivative works, without royalty or limitation,
 the subject matter of the dominating interest.  This license provision
 will be binding on the assignees of, or other successors to, the
 dominating interest, as well as on Developer.  This license provision
 is nonexclusive, royalty-free and non-cancellable.
 
 7. Developer warrants to FSF that its reports under Section 2 above
 are accurate and that it is the sole copyright holder of all changes
 and/or enhancements submitted to the FSF under this agreement, as made
 by or otherwise acquired by Developer.
 
 Developer agrees to hold FSF harmless from damage arising out of any
 breach of this warranty---any adverse ownership interest not reported
 by Developer and resulting in direct or indirect harm to FSF.
 Developer's warranty and hold harmless obligation does not run to the
 general public except insofar as damage to the public results in harm
 to FSF.
 
 Developer is not obliged to defend FSF against any spurious claim of
 adverse ownership, but will cooperate with FSF in defending against
 any such claim and in any event be responsible if the claim is not
 spurious, i.e. if the adverse claimant is a true owner and FSF is
 thereby damaged.
 
 Developer makes no other express or implied warranty (including
 without limitation, in this disclaimer of warranty, any warranty of
 MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE).
 
 8. This agreement and all conveyances made pursuant hereto are deemed
 made in Massachusetts, this agreement and all such conveyances being
 deemed made under seal, as of the date of fixation of any work covered
 hereby in tangible form and, also, as of the date of execution of this
 agreement by Developer.
 
          Developer (organization):\hrulefill,
 
         
 	 Signature:\hrulefill
 
 
          Name of signer:\hrulefill
 
 
          Title:\hrulefill
 
 
          Date:\hrulefill
 
 
 ACCEPTED: FREE SOFTWARE FOUNDATION
 
 Richard Stallman, President

\newpage
\centering DESCRIPTION OF WORKS
 
\vspace{2cm}

\begin{center}
\large 
{\bfseries The DROOPI (Distributed Reusable Object-Oriented
Polymorphic Infrastructure) middleware components}\\
A library of runtime components for the implementation
of object-oriented distributed systems, and associated
tools. These library and tools constitute a contribution
to the GNAT compiler and runtime library.
\end{center}

%%  \begin{tabular}{lcc}
%%    & Title of File     &        Date(s) of Developer's Work Only \\
%%  &&\\
%%   1.& \hrulefill   &    \hrulefill  \\
%%   2.& \hrulefill   &    \hrulefill  \\
%%   3.& \hrulefill   &    \hrulefill  \\
%%   4.& \hrulefill   &    \hrulefill  \\
%%   5.& \hrulefill   &    \hrulefill
%%  \end{tabular} 
 
%%                                    _ _ _ _ _ _ _ _ _ _ _ _ _ _ _
%%                                    *Includes Developer's work and
%%                                    work of others for Developer 
%%                                    [see par. 2, above].  If 
%%                                    others are involved, please
%%                                    indentify them and briefly 
%%                                    state the manner of their 
%%                                    conveyance of rights --- Work
%%                                    for hire? Assignment? To 
%%                                    Developer? Direct to FSF? 
%%                                    - Thank you.


\end{document}
